\jaxcosmo: An End-to-End Differentiable and GPU Accelerated Cosmology Library

### 2022-11-15 paper_OJA.tex

Here is the simple way \numpyro\ deals with SVI initialisation and optimisation\footnote{See also listing in Appendix \ref{app-DESY1}) where the \texttt{model} is described and the \texttt{cl\_obs} are the observed $C_\ell$.}
\begin{lstlisting}[language=iPython]
import numpyro.infer.autoguide as autoguide
from numpyro.infer import SVI, Trace_ELBO
from numpyro.optim import Adam
# choose a SVI guide 
guide = autoguide.AutoMultivariateNormal(model,
        init_loc_fn=numpyro.infer.init_to_median())
# SVI optimisation 
optimizer = numpyro.optim.Adam(1e-3)
svi = SVI(model_spl, guide,optimizer,loss=Trace_ELBO(num_particles=10))
svi_result = svi.run(jax.random.PRNGKey(0),20_000, cl_obs)
#To get samples from SVI approximated posterior 
samples = guide.sample_posterior(jax.random.PRNGKey(1), svi_result.params, sample_shape=(100_000,))
\end{lstlisting}
%
\JEC{Once the optimisation is done on gets $z$ i.i.d. samples from the $q(z,\lambda^\ast)$ distribution  applying the inverse of the inverse of the $S_{\lambda^\ast}$ and $T$ transformations. Using the same DES-Y1 simulation as in previous section, we use both a MultiVariate Normal distribution (MVN) and a Block Nueral Autoregressive Flow (B-NAF) \citep{pmlr-v115-de-cao20a} as \textit{guides} to approximate the true posterior (\textit{model}). The reader can have a glance to the following notebook \nblink{DESY_Y1_SVI_and_NeutraReparam}. Concerning the B-NAF architecture it is composed of  a single flow using a block autoregressive structure with 2 hidden layers of 8 units each. The SVI optimization has been performed with the Adam optimizer \citep{2014arXiv1412.6980K} and a learning rate set to $10^{-3}$. We have stopped the optimization after 20,000 (30,000) steps to ensure a stable ELBO loss convergence when using the B-NAF (MVN) guides, and no tuning of the optimizer has been undertaken as for instance a learning rate scheduling. It takes about 2 or 3 hours on the NVidia V100 GPU scaling according to the number of steps}.  \FrL{Arf... 2 hours !}

@@

The \textit{Neural Transport} is accessible throw the  \verb|NeuTraReparam| in \numpyro\ library as followed (See the notebook \nblink{DESY_Y1_SVI_and_NeutraReparam}):

@@

mcmc.run(jax.random.PRNGKey(42))
\end{lstlisting}
%\end{minted}
More details can be found in the repository of this paper to perform a \numpyro\ modeling using \jaxcosmo : a structured NUTS sampling described in Section~\ref{sec-NUTS} \nblink{VectorizedNumPyro_block}, and a SVI optimisation followed by NUTS sampling of Section~\ref{sec-Neural-Reparametrisation} \nblink{DESY_Y1_shear_cluster_ModelReparam_SVI-MVN_Neutra_NUTS}. Notice however that the notebooks are only illustrative and for a real sample production the authors invite the reader to run their own script.

>>>>

Here is the simple way \numpyro\ deals with SVI initialisation and optimisation\footnote{See also listing in Appendix \ref{app-DESY1}) where the \texttt{model} is described and the \texttt{cl\_obs} are the observed $C_\ell$.}
\begin{lstlisting}[language=iPython]
import numpyro.infer.autoguide as autoguide
from numpyro.infer import SVI, Trace_ELBO
from numpyro.optim import Adam
# choose a SVI guide 
guide = autoguide.AutoMultivariateNormal(model,
        init_loc_fn=numpyro.infer.init_to_median())
# SVI optimisation 
optimizer = numpyro.optim.Adam(1e-3)
svi = SVI(model_spl, guide,optimizer,loss=Trace_ELBO(num_particles=10))
svi_result = svi.run(jax.random.PRNGKey(0),20_000, cl_obs)
#To get samples from SVI approximated posterior 
samples = guide.sample_posterior(jax.random.PRNGKey(1), svi_result.params, sample_shape=(100_000,))
\end{lstlisting}
%
\JEC{Once the optimisation is done on gets $z$ i.i.d. samples from the $q(z,\lambda^\ast)$ distribution  applying the inverse of the inverse of the $S_{\lambda^\ast}$ and $T$ transformations. Using the same DES-Y1 simulation as in previous section, we use both a MultiVariate Normal distribution (MVN) and a Block Nueral Autoregressive Flow (B-NAF) \citep{pmlr-v115-de-cao20a} as \textit{guides} to approximate the true posterior (\textit{model}). The reader can have a glance to the following notebook \nblink{DES_Y1_SVI_and_NeutraReparam}. Concerning the B-NAF architecture it is composed of  a single flow using a block autoregressive structure with 2 hidden layers of 8 units each. The SVI optimization has been performed with the Adam optimizer \citep{2014arXiv1412.6980K} and a learning rate set to $10^{-3}$. We have stopped the optimization after 20,000 (30,000) steps to ensure a stable ELBO loss convergence when using the B-NAF (MVN) guides, and no tuning of the optimizer has been undertaken as for instance a learning rate scheduling. It takes about 2 or 3 hours on the NVidia V100 GPU scaling according to the number of steps}.  \FrL{Arf... 2 hours !}

@@

The \textit{Neural Transport} is accessible throw the  \verb|NeuTraReparam| in \numpyro\ library as followed (See the notebook \nblink{DES_Y1_SVI_and_NeutraReparam}):

@@

mcmc.run(jax.random.PRNGKey(42))
\end{lstlisting}
%\end{minted}
More details can be found in the repository of this paper to perform a \numpyro\ modeling using \jaxcosmo : a structured NUTS sampling described in Section~\ref{sec-NUTS} \nblink{VectorizedNumPyro_block}, and a SVI optimisation followed by NUTS sampling of Section~\ref{sec-Neural-Reparametrisation} \nblink{DESY_Y1_SVI_and_NeutraReparam}. Notice however that the notebooks are only illustrative and for a real sample production the authors invite the reader to run their own script.
